% !TEX program = xelatex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Title: SOP LaTeX Template
%% Author: Soonho Kong / soonhok@cs.cmu.edu
%% Modified by: Eric Silk
%% Created: 2012-11-12
%% Modified: 2021
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% Requirement:
%%     You need to have the `Adobe Caslon Pro` font family.
%%     For more information, please visit:
%%     http://store1.adobe.com/cfusion/store/html/index.cfm?store=OLS-US&event=displayFontPackage&code=1712
%%
%% How to Compile:
%%     $ xelatex main.tex
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[letterpaper]{article}
\usepackage[letterpaper,margin=1.0in,noheadfoot]{geometry}
\usepackage{fontspec, color, enumerate, sectsty}
\usepackage[normalem]{ulem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                      YOUR INFORMATION
%
%      PLEASE EDIT THE FOLLOWING LINES ACCORDINGLY!!
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\soptitle}{Statement of Purpose}
\newcommand{\yourname}{Eric Silk}
\newcommand{\youremail}{esilk16@uw.edu}

%% FONTS SETUP
\defaultfontfeatures{Mapping=tex-text}
\usepackage[bookmarks, colorlinks, breaklinks,
pdftitle={\yourname - \soptitle},pdfauthor={\yourname}, unicode]{hyperref}
\hypersetup{linkcolor=magneta,citecolor=magenta,filecolor=magenta,urlcolor=[named]{WildStrawberry}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                      Title and Author Name
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\begin{center}{\huge \scshape \soptitle}\end{center}
\begin{center}\vspace{0.2em} {\Large \yourname\\}
  {\youremail\\}{CSEM PhD\\}{EID eas5573}\end{center}
\frenchspacing



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                      SOP Body
% NOTE: Use \amper instead of \&
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Motivation and Background}
I am a Research Engineer at Schweitzer Engineering Laboratories, Inc., and graduating masters
student of Applied Mathematics at the University of Washington with a Bachelors of Science
in Electrical Engineering from the University of Idaho. My employer's mission is
one I believe in: ``To make electric power safer, more reliable, and more economical.'' On a daily
basis I perform at the intersection of applied math, electrical engineering, and software
engineering to further this goal.

Upon transferring to SEL's Government Services division from R\&D, it became apparent
there is a need for people who are capable of bridging the gap between the abstract
and esoteric world of mathematics and real-world applications in engineering.
For this reason, I elected to pursue graduate education.
Even with the huge strides I have made, my work has just started. Earning a PhD will allow me to
transition into an independent, productive researcher and to speak with authority on topics
related to electric power.

\section*{Relevant Experience}
\paragraph{Career Research}
Due to the nature of my work for my employer, I am not allowed to discuss many specifics
and definitely not allowed to publish. We are recognized as experts
on power systems by various federal agencies and regularly provide research, analysis,
and product in this space. In particular, the Data Analytics group marries this expertise
with high performance computing, machine learning, and traditional applied mathematics to
achieve novel and challenging goals. Of what I can discuss, the following stand out.

In February, I submitted a patent application for a compression method
to help streamline the massive data output of SEL's flagship relay, the T400L/T401L --- an
18-bit, megahertz sampling relay for the detection of traveling wave events.
We have a vested interest in long-term data captures at this
fidelity, but the existing implementation had excessive storage requirements,
(four terabytes for three weeks). During Dr. Kutz's class (AMATH 582), I applied his theme of
assuming structure to solve $Ax=b$ given only $b$ and
assume an autoregressive model and produce lossy compression. Then, the errors are stored in
Golomb-Rice codes, producing a reduction of 5.7 to 7.4 times less than prior art.
This was a major achievement --- it enabled storing one year of data on a single 3.5"
drive, or streaming using less bandwidth than Netflix.

My first assignment in Government Services was developing sensor system to measure and subsequently
store Synchrophasor data (GPS timestamped frequency, voltage, and phase) on the cloud. This involved
selecting a computing platform, developing code to interact with a phasor measurement unit (PMU),
robust logging and system restarts, and code for Amazon Web Services' Internet of Things,
Lambda, and S3 services -- all while ensuring the devices were secure. We successfully deployed
several devices that have now performed uninterrupted for several years.


A recent achievement was the synthesis of a massive schematic dataset for bootstrapping a machine
learning model. Over ten-thousand Eagle schematics were generated, including components such as
passives (resistors, etc.), active components (diodes, BJT's), integrated circuits, and more.
This included bounding boxes and component identifiers --- all without human intervention. I am
actively expanding this capability to include more CAD programs.

Finally, I have applied optimization in several influential ways. The first was generating
nonlinear transmission lines via genetic algorithms to achieve particular performance
characteristics that would be intractable with traditional engineering techniques.
From this, two lessons were gleaned: have frank discussions about
requirements minimize iterations; and, while open-source software is a boon, niche projects can
suffer from instability, end-of-life components, or even incorrect behaviors.
Another project involved fitting high-order, nonlinear transformer parameters through genetic
algorithms. This was quite successful, taught me that judicious use of penalties prevents
computationally correct but physically nonsensical models, and is still in use by our power
engineering team. Finally, I implemented a crest-factor minimization problem in PyTorch that will
serve as a testbed for different optimizers, and introduced autodifferentiation and quasi-newton
methods into our repertoire.


\paragraph{Independent Study and Scalable Second Order Optimizers}
As an online Masters student, I was unable to write a thesis. To ameliorate this, I arranged an
independent study course with Dr. Andrew Lumsdaine.
Topics ranged from power system simulations using PETSc, matrix decompositions in CUDA, and the
intersection of graphs and linear algebra. I learned an incredible amount and made enough of an
impression to help with his Second Order Methods for Scalable Optimization project. This project
hinges around the use of finite-difference Hessian approximations, along with limited-memory
techniques to accelerate machine learning training. Our git repository can be found here:
https://github.com/lums658/ml20

\section*{Career Goals}
I foresee three career trajectories: professorship, industry, and civil service.
All share a common goal: perform at the intersection of applied math and engineering
and provide relevant, timely, and digestible research and education to students,
peers, or customers.

As an educator, I would provide power engineers with the tools
to step beyond traditional power engineering analyses and engage with cutting-edge applied
mathematics techniques, while providing context and constraints to keep these techniques grounded in
industrial realities.  In industry or civil service, I would continue in much my same role:
exemplary work for government customers making the grid safer, cheaper, and more reliable.
This would take the form of consulting and providing tools for use by the federal
government, and sound advice for shaping policy.

\section*{Challenges}
I want to address two points. First, I had a medical withdrawal in undergrad with lingering
effects in subsequent semesters. This issue has since been fully resolved, as evidenced
by my final semesters and my graduate performance. Second, my withdrawal in graduate school was
due to a death in my family. Neither pose risks to future academic endeavors.

\section*{Institution Selection}
When I asked Dr. Lumsdaine his recommendation for programs that would meet my research and career
goals, he stated without hesitation: ``UT Austin." I trust his advice implicitly, but have additionally
had the pleasure of meeting two other graduates of your undergrad program (one in my current
program and another at a conference) who echoed his sentiment. All evidence suggests a highly
collaborative environment, with options for both Electrical Engineering and Computational Engineering
for coursework and advising. This is exactly where I hope to conduct research -- rigorous mathematics
and clever algorithms supporting meaningful and far-reaching engineering to support the next
generation of power systems design.

\section*{Faculty Selection}
It was requested to include which faculty I would be interested in working with. Without having
had the opportunity to meet with any of them and assess fit, I cannot reliably make a
determination on particular individuals. However, I do have confidence there will be someone
with whom I can establish a mutually beneficial research relationship.


\section*{Other Institutions Applied To}
As requested by the instructions, the following section details the other programs I have
applied to and my rationale for such.

\subsection*{University of Texas, Austin: ECE}
I have applied to the ECE program at UT Austin as well as CSEM. While I hope my reasoning
has been explained so far, I want to straddle CSEM and ECE. Either department seems
to represent the opportunity to do so, provided I'm able to find co-advisors or committee
members amenable to this arrangement.

\subsection*{Georgia Institute of Technology: ECE and CSE}
GA Tech represents one of the other major programs with an exceptionally strong computational
science and mathematics program, as well as an excellent ECE program -- I've applied to both.
For this reason, it was the second suggestion of my advisor following UT Austin.
Additionally, an information session I attended hosted by a professor in CSE confirmed the
ability to collaborate with professors within either department, provided I'm admitted to at
least one.

\subsection*{University of Washington: ECE and AMATH}
I am currently a student at the UW, and have been highly impressed with the faculty and my
peers within the AMATH program. It seems obvious to continue in this environment. Several
professors, including Dr. Kutz, are the experts in sparse system identification techniques
that I suspect will have powerful applications in power systems analysis. In addition,
there are locale considerations, as well as the opportunity for a joint UW-PNNL fellowship
opportunity.

\subsection*{Purdue: ECE}
Purdue has one of the best power engineering programs in the country. My employer maintains
strong relationships with this university, going so far as to have endowed a professorship
and to construct new manufacturing and engineering facilities a stone's throw from campus.
This presents an opportunity for continued collaboration with my employer, and a trustworthy
endorsement of the merits of the program. Additionally, they offer a non-degree option
for computational science and engineering that one can apply for after admission into a
degree seeking program --- I intend to pursue this, should I be admitted and elect to attend.

\subsection*{University of Illinois, Urbana-Champaign: ECE and CSE}
It seems half the papers I've read related to high-performance computing have been
authored, at least in part, by an affiliate of UIllinois. This was also a top
recommendation from my advisor if I were interested in emphasizing more in HPC
(a notion I'm amenable to). This is another school offering both ECE and CSE.

\subsection*{Oregon State University: EE}
I've had the privilege to collaborate briefly with several students and faculty from OSU's
EE program on a project sponsored by the Department of Energy. I was impressed with their
contributions, and the atmosphere the advising professors fostered, and they seemed to
be practiced at finding industry collaborators to ground their research. Similar to the UW,
they also have a joint PNNL fellowship that is highly alluring.

\subsection*{Massachusetts Institute of Technology: EECS}
At risk of sounding flippant, this program needs little explanation. I've now been able to
work with several faculty who graduated from here (Dr. Andrew Lumsdaine, Dr. Tony Chiang)
as well as being an avid subscriber of their open source courseware. In fact, I was a
pilot student for the first iteration of what is now edX (originally MITx) in their
6.002x course during my senior year of high school --- I even managed a B! Every time,
I've been awed by the ability of their graduates to develop powerful abstractions for
complex topics and disseminate them in a way that makes it obvious.

\section*{Concluding Remarks}
I hope I've demonstrated a clear upward trajectory in my career and research.
I'm on the cusp of being a strong and independent researcher; a PhD from your program
would be the keystone in my ability to contribute to electrical engineering and applied
mathematics.
\end{document}
